---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# HealthDataScotland

<!-- badges: start -->
[![R-CMD-check](https://github.com/jsleight1/HealthDataScotland/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/jsleight1/HealthDataScotland/actions/workflows/check-standard.yaml)
[![Codecov test coverage](https://codecov.io/gh/jsleight1/HealthDataScotland/branch/main/graph/badge.svg)](https://app.codecov.io/gh/jsleight1/HealthDataScotland/?branch=main)
<!-- badges: end -->

This repository contains an R package for the HealthDataScotland shiny application.
This shiny application downloads, processes and presents demographic and
bed capacity data related to GP practices and hospitals across Scotland, respectively.

## Installation

You can install HealthDataScotland from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
devtools::install_github("jsleight1/HealthDataScotland")
```

Please refer to the [documentation](https://jsleight1.github.io/HealthDataScotland/) for further details regarding the design
and usage of this package.
