---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# HealthDataScotland

<!-- badges: start -->
[![R-CMD-check](https://github.com/jsleight1/HealthDataScotland/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/jsleight1/HealthDataScotland/actions/workflows/check-standard.yaml)
[![Codecov test coverage](https://codecov.io/gh/jsleight1/HealthDataScotland/branch/1-development/graph/badge.svg)](https://app.codecov.io/gh/jsleight1/HealthDataScotland/?branch=1-development)
<!-- badges: end -->


HealthDataScotland contains a shiny application for the presentation 
of general practice (GP) demographic and hospital bed capacity data. The shiny application presents interactive maps of the locations 
of GP practices and hospital across Scotland. Indiviual GP practices and hospitals can be clicked to interogate the data was single centre. Conversely comparisons can be made across centres by using the map draw function to selected multiple health centres. The processed data is also presented in a table format. This can be downloaded in full or can be filtered to download a subset of the data set.

## Installation

You can install the development version of HealthDataScotland from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("jsleight1/HealthDataScotland")
```

You can also install HealthDataScotland using [renv](https://rstudio.github.io/renv/articles/renv.html) to create a reproducible environment for development.

## Example

The shiny application can be launched locally using:

```{r, eval = FALSE}
library(HealthDataScotland)
health_data_scotland()
```

## Data and references.

GP and Hospital is downloaded from the [Scottish Health and Social Care Open Data
platform](https://www.opendata.nhs.scot/) using the
[phsopendata](https://github.com/Public-Health-Scotland/phsopendata/) R package. 
Spatial data for the generation of interactive maps are obtained 
from the [Spatial Hub Scotland](https://data.spatialhub.scot/) and from data.gov.uk. For further information please refer to these databases.

