---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# HealthDataScotland

<!-- badges: start -->
[![R-CMD-check](https://github.com/jsleight1/HealthDataScotland/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/jsleight1/HealthDataScotland/actions/workflows/check-standard.yaml)
[![Codecov test coverage](https://codecov.io/gh/jsleight1/HealthDataScotland/branch/main/graph/badge.svg)](https://app.codecov.io/gh/jsleight1/HealthDataScotland/?branch=main)
<!-- badges: end -->

This repository contains an R package for the [HealthDataScotland](https://www.healthdatascotland.co.uk/){target="_blank"} shiny application.
This shiny application downloads, processes and presents demographic and
bed capacity data related to GP practices and hospitals across Scotland, respectively.

## Installation

You can install HealthDataScotland from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
devtools::install_github("jsleight1/HealthDataScotland")
```

## Examples

### R6 objects

```{r, eval = FALSE}
gp_unit <- example_gp_unit()
gp_unit$metadata()
gp_unit$plot(type = "population_pyramid")
```

```{r, eval = FALSE}
hospital_unit <- example_hospital_unit()
hospital_unit$metadata()
hospital_unit$plot(type = "specialty_line")
```

```{r, eval = FALSE}
map_unit <- example_map_unit()
map_unit$plot(type = "interactive_map")
```

### Local shiny usage

```{r, eval = FALSE}
Sys.setenv("R_CONFIG_ACTIVE" = "development")
save_processed_data()
health_data_scotland()
```

Please refer to the [documentation](https://jsleight1.github.io/HealthDataScotland/)
for further details regarding the design and usage of this package.


# Disclaimer

This R package is licensed using a GNU General Public License and contains
data from a variety of sources licensed under the Open Government License (OGL).
As this application is primarily a hobby project, the author is likely to make
no financial gain from this application and therefore accepts no liability,
responsibility or guarantees as to how the package is designed or how the
package downloads, processes and presents the data. The original
published data sets should always be consulted when using this application.
